Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-21 alloscore_manuscript/supplement_short.tex
minor suggested changes to supplement.

Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined 
by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.

A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other forecast $G$ 
as being better than $F$ itself, that is, that
\begin{align}
\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)] \label{eqn:prop_ineq}
\end{align} 
for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of 
a forecaster focussed (solely) on expected loss minimization, the decision to state 
a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$. 
$S$ is \emph{strictly proper} when 
\eqref{eqn:prop_ineq} is a strict inequality, in which case the 
\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.

@@

Here we study the form of the Bayes act for the allocation problem (AP) (\eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}] 
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
is a vector of quantiles of the marginal forecast distributions $F_i$ at a single probability level
$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

@@

We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$
is an increasing function of $\tau$.
The algorithm begins with an initial search interval $[\tau_L, \tau_U]$ that is known to contain the solution $\tau^{\star}$.
This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
At each step of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M},F_i}$ at the midpoint of the search interval, $\tau_M = \frac{1}{2}(\tau_L + \tau_U)$. We then narrow the search to the sub-interval that contains the solution:
\begin{itemize}
  \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
  \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
\end{itemize}
This search continues until a stopping condition, defined in terms of the width of the search interval, is met.
We have implemented this procedure along with the resulting score computations in the 
R package \verb`alloscore` which provided all allocation score values used in the analysis of section \ref{sec:application} 
in the main text.

% Subtleties can arise when the forecast
% densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level, 
% leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$, 
% \eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
% We have adopted conventions for detecting such levels and enforcing consistency in score calculations
% near them. Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.

@@

In section 3 of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the COVID-19 Forecast Hub. These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$. For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R (cite). This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.

We refer the reader to the \verb`distfromq` documentation for further detail.

>>>>

Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined
by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.

A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
as being better than $F$ itself, that is, that
\begin{align}
\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)] \label{eqn:prop_ineq}
\end{align}
for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of
a forecaster focussed (solely) on expected loss minimization, the decision to state
a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$.
$S$ is \emph{strictly proper} when
\eqref{eqn:prop_ineq} is a strict inequality, in which case the
\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.

@@

Here we study the form of the Bayes act for the allocation problem (AP) (equation \eqref{eqn:loss_fn} in section \ref{sec:methods.detailed.specific_allocation}) of the text:
\begin{align}
    \underset{0 \leq x}{\mathrm{minimize}}\,\, \mathbb{E}_{F} [s_A(x, Y)]= \sum_{i=1}^{N} L \cdot \mathbb{E}_{F_i}[(Y_i - x_i)_{+}]
     \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
where the marginal forecasts $F_i$ for $i=1,\dots,N$ represent forecasts for $N$ distinct locations.
We show that the Bayes act $x^{F,K} = (x_1^{F,K},\ldots,x_N^{F,K})$ for a forecast $F$ and resource constraint level $K$
is a vector of quantiles of the marginal forecast distributions $F_i$ at a single probability level
$\tau^{F,K}$, that is, $x_i^{F,K} = q_{F_i,\tau^{F,K}}$. An immediate consequence used in the examples in Section \ref
{sec:methods.overview} in the main text is that if $F_i = \mathrm{Exp}(1/\sigma_i)$ for all $i$, then the Bayes act is
proportional to $(\sigma_1,\ldots,\sigma_N)$, since $q_{\mathrm{Exp}(1/\sigma),\tau} = -\sigma \log(1-\tau)$.

@@

We have implemented an iterative bisection method that makes use of the fact that $\sum_{i=1}^N q_{\tau,F_i}$
is an increasing function of $\tau$.
The algorithm begins with an initial search interval $[\tau_{L,1}, \tau_{U,1}]$ (such as $[0,\max_{i}F_i(K)]$) that clearly contains
the solution $\tau^{\star}$.
% This can be confirmed by verifying that the resource allocations corresponding to the lower endpoint of the search interval do not exceed the resource constraint ($\sum_{i=1}^N q_{\tau_L,F_i} \leq K$) and the resource allocations corresponding to the upper endpoint match or exceed the available resources ($\sum_{i=1}^N q_{\tau_U,F_i} \geq K$).
At each step $j$ of the algorithm, we evaluate the total allocation $\sum_{i=1}^N q_{\tau_{M,j},F_i}$ at the midpoint of the search interval,
$\tau_{M,j} = \frac{1}{2}(\tau_{L,j} + \tau_{U,j})$ and continue the search on the narrowed sub-interval
\begin{align}
[\tau_{L,j+1},\tau_{U,j+1}] =
\begin{cases}
[\tau_{L,j}, \tau_{M,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} \geq K \\
[\tau_{M,j}, \tau_{U,j}] & \text{if } \sum_{i=1}^N q_{\tau_{M,j},F_i} < K.
\end{cases} \nonumber
\end{align}
% We then narrow the search to the sub-interval that contains the solution:
% \begin{itemize}
%   \item If $\sum_{i=1}^N q_{\tau_{M},F_i} \geq K$, we continue the search on the interval $[\tau_L, \tau_M]$.
%   \item If $\sum_{i=1}^N q_{\tau_M, F_i} \leq K$, we continue the search on the interval $[\tau_M, \tau_U]$.
% \end{itemize}
This search continues until $\tau_{U,j+1} < (1+\varepsilon)\tau_{L,j+1}$ for a suitably small $\varepsilon>0$.
We have implemented this procedure along with the resulting score computations in the
R package \verb`alloscore` which provided all allocation score values used in the analysis of section \ref{sec:application}
in the main text.

Subtleties can arise when the forecast
densities $f_i$ vanish or are very small, in which case quantiles are non-unique or highly variable near a probability level,
leading to ambiguity or numerical instabilities in the evaluation of $\sum_{i=1}^N q_{\tau,F_i}$. Additionally, if point masses are present in any of the $F_i$,
\eqref{eqn:quantiles-sum-to-K} will not have a unique solution for some discrete set of constraint levels $K$.
We have adopted conventions for detecting such levels and enforcing consistency in score calculations
near them. Through extensive experimentation, we have determined that these conditions seem to address these challenges with the forecasts we are working with, but we leave a more rigorous approximation error analysis for later work.

@@

In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19 hospitalizations that have been submitted to the US COVID-19 Forecast Hub. These forecasts are submitted to the Hub using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01, 0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2, \allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$. For a given week and target date, we use $q_{i,k}$ to denote the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq} This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that matches the two most extreme quantiles in each tail.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.

We refer the reader to the \verb`distfromq` documentation for further detail.\citep{ray-distfromq}
